---
title: "fish_dt_CAFP"
author: "Stijn Bruneel"
date: '2022-10-05'
output: word_document
---

```{r setup, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
# import libraries from file
source("./source/not_functions/libraries.R")
```

```{r}
source("./source/functions/log_model_state.R")
source("./source/functions/plot_histogram_length_weight_distribution.R")
source("./source/functions/run_chi_squared.R")
source("./source/functions/stats_data_level1.R")
source("./source/functions/stats_data_level2a.R")
source("./source/functions/rmse.R")
```

```{r}
df_OK_f <- read.csv("./data/fish/internal/df_OK_f.csv")
df_OK_n_ad <- read.csv("./data/fish/internal/df_OK_n_ad.csv")
df_OK_f <- bind_rows(df_OK_f,df_OK_n_ad)
df_OK_f <- df_OK_f[which(df_OK_f$Species=="roach"),]

df_OK_f$Frequency<-as.factor(df_OK_f$Frequency)
df_OK_f$rpm<-as.factor(df_OK_f$rpm)

Language="english"

df_OK_f$nState=NA
df_OK_f$nState[which(df_OK_f$State=="Alive")]=1
df_OK_f$nState[which(df_OK_f$State=="Dead")]=0

df_OK_f$State=as.factor(df_OK_f$State)
df_OK_f$State <- relevel(df_OK_f$State, ref = "Dead")
```

```{r}
df_OK_f_long_type <- read.csv("./data/fish/internal/df_OK_f_long_type.csv")
df_OK_n_ad_long_type <- read.csv("./data/fish/internal/df_OK_n_ad_long_type.csv")
df_OK_f_long_type <- bind_rows(df_OK_f_long_type,df_OK_n_ad_long_type)
df_OK_f_long_type<- df_OK_f_long_type[which(df_OK_f_long_type$Species=="roach"),]

df_OK_f_long_type<-df_OK_f_long_type[-which(df_OK_f_long_type$Value==0 | df_OK_f_long_type$Injury=="Injury_type_dead"),]
if (exists("Language")){
  if (Language=="dutch"){
    df_OK_f_long_type$Injury[which(df_OK_f_long_type$Injury=="Injury_type_no")] = "geen"
    df_OK_f_long_type["Injury"][df_OK_f_long_type["Injury"] == "Injury_type_no"] <- "geen"
    df_OK_f_long_type$Injury[which(df_OK_f_long_type$Injury=="Injury_type_slight")] <- "licht"
    df_OK_f_long_type$Injury[which(df_OK_f_long_type$Injury=="Injury_type_severe")] <- "zwaar"
  }
  if (Language=="english"){
    df_OK_f_long_type$Injury[which(df_OK_f_long_type$Injury=="Injury_type_no")] = "no"
    df_OK_f_long_type["Injury"][df_OK_f_long_type["Injury"] == "Injury_type_no"] <- "no"
    df_OK_f_long_type$Injury[which(df_OK_f_long_type$Injury=="Injury_type_slight")] <- "slight"
    df_OK_f_long_type$Injury[which(df_OK_f_long_type$Injury=="Injury_type_severe")] <- "severe"
    df_OK_f_long_type$Injury <- factor(df_OK_f_long_type$Injury, levels=c("no", "slight", "severe"))
  }
}
df_OK_f_long_type$Injury<-as.factor(df_OK_f_long_type$Injury)
df_OK_f_long_type$Control<-as.factor(df_OK_f_long_type$Control)
df_OK_f_long_type$Frequency<-as.factor(df_OK_f_long_type$Frequency)
df_OK_f_long_type$frpm <- as.factor(df_OK_f_long_type$rpm)
df_OK_f_long <- read.csv("./data/fish/internal/df_OK_f_long.csv")
df_OK_n_ad_long <- read.csv("./data/fish/internal/df_OK_n_ad_long.csv")
df_OK_f_long <- bind_rows(df_OK_f_long,df_OK_n_ad_long)
df_OK_f_long<-df_OK_f_long[-which(df_OK_f_long$Value==0 | df_OK_f_long$Injury=="Class_4"),]
df_OK_f_long$Injury<-as.factor(df_OK_f_long$Injury)
df_OK_f_long$Control<-as.factor(df_OK_f_long$Control)
df_OK_f_long$Frequency<-as.factor(df_OK_f_long$Frequency)
df_OK_f_long$frpm <- as.factor(df_OK_f_long$rpm)
```

```{r}
specieslist_OK <- "roach"
injurylist_OK <- read.csv("./data/fish/internal/injurylist_OK.csv")[,2]
injurylist_type_OK <- read.csv("./data/fish/internal/injurylist_type_OK.csv")[,2]
```

```{r}
df_OK_f$Type=df_OK_f$Control
df_OK_f_long_type$Type=df_OK_f_long_type$Control
df_OK_f_long$Type=df_OK_f_long$Control
```

## Methodology

In addition to the logistic models, to assess the survival probability, and multinomial models, to assess the injury probability, decision trees were constructed. To assess the survival rate for the different scenarios of the FNAFP, decision trees were developed for the entire dataset and for each species separately. These provide an intuitive and clear overview of the data and underlying patterns. The state of each fish (alive=1 versus dead=0) was used as response. The explanatory variables were species (discrete), rpm (discrete) and fish length (continuous).

## Results


```{r}
# CRT for injury OK
control <- rpart.control(minsplit = 10)
color.selection <- "GnYlRd"
par(mfrow = c(1,1))

pdf("./figures/dt_injury_type_all_natural.pdf")
form.state.df.tree <- as.formula(Injury ~ Pump_type_rpm + Type + Length)
model.injury.df.tree <- rpart(form.state.df.tree, data = df_OK_f_long,control=control)
e<-fancyRpartPlot(model.injury.df.tree,main="roach",box.palette = color.selection, legend.x = NA, palettes = "PuRd", xflip=TRUE)
e
model.injury.df.tree$variable.importance/max(model.injury.df.tree$variable.importance)
dev.off()
```

No decision tree could be constructed for the CAFP

```{r}
# Random forest state OK
form.state.df.forest <- as.formula(nState ~ Pump_type_rpm + Species + Length + Control)
model.state.df.forest <- ranger(form.state.df.forest,df_OK_f[which(is.na(df_OK_f$Length)==FALSE),],num.trees=500,respect.unordered.factors = "order",importance="permutation")
model.state.df.forest$variable.importance
rmse(predict(model.state.df.forest,df_OK_f[which(is.na(df_OK_f$Length)==FALSE),])$predictions,df_OK_f[which(is.na(df_OK_f$Length)==FALSE),])

form.state.df.forest <- as.formula(nState ~ Pump_type_rpm + Species + Length)
model.state.df.forest <- ranger(form.state.df.forest,df_OK_f[which(is.na(df_OK_f$Length & df_OK_f$Control=="pump")==FALSE),],num.trees=100,respect.unordered.factors = "order",importance="permutation")
model.state.df.forest$variable.importance
rmse(predict(model.state.df.forest,df_OK_f[which(is.na(df_OK_f$Length & df_OK_f$Control=="pump")==FALSE),])$predictions,df_OK_f[which(is.na(df_OK_f$Length & df_OK_f$Control=="pump")==FALSE),])

k=0
model.state.df.forest<-list()
for(i in specieslist_OK){
  print(i)
  k=k+1
  form.state.df.forest <- as.formula(nState ~ Pump_type_rpm + Length + Control)
  model.state.df.forest[[k]] <- ranger(form.state.df.forest,df_OK_f[which(df_OK_f$Species==i & is.na(df_OK_f$Length)==FALSE),],num.trees=100,respect.unordered.factors = "order",importance="permutation")
  print(model.state.df.forest[[k]]$variable.importance)
  print(rmse(predict(model.state.df.forest[[k]],df_OK_f[which(df_OK_f$Species==i & is.na(df_OK_f$Length)==FALSE),])$predictions,df_OK_f[which(df_OK_f$Species==i & is.na(df_OK_f$Length)==FALSE),]))

}

k=0
model.state.df.forest<-list()
for(i in specieslist_OK){
  print(i)
  k=k+1
  form.state.df.forest <- as.formula(nState ~ Pump_type_rpm + Length)
  model.state.df.forest[[k]] <- ranger(form.state.df.forest,df_OK_f[which(df_OK_f$Species==i & is.na(df_OK_f$Length)==FALSE & df_OK_f$Control=="pump"),],num.trees=100,respect.unordered.factors = "order",importance="permutation")
  print(model.state.df.forest[[k]]$variable.importance)
  print(rmse(predict(model.state.df.forest[[k]],df_OK_f[which(df_OK_f$Species==i & is.na(df_OK_f$Length)==FALSE & df_OK_f$Control=="pump"),])$predictions,df_OK_f[which(df_OK_f$Species==i & is.na(df_OK_f$Length)==FALSE & df_OK_f$Control=="pump"),]))

}
```

